### 2020-11-11 latex/acrobot.tex
Edits to Manfredi's improvements of VLP proof
Changing regular ref to eqref for equations

Since we are now working with a simple acrobot, 
we have \(l_{c_u} = l_{c_a} = l_u = l_a = l\) and
\(m_u = m_a = m\). 
On top of this, the moments of inertia \(J_u\) and \(J_a\) of the rods vanish.
Reducing
(\ref{eqn:general-acrobot-inertia}-\ref{eqn:general-acrobot-potential})
yields the simplified inertia matrix \(D_s\) and potential function \(P_s\),
where
\begin{align}
    D_s(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    P_s(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    .
\end{align}
\begin{notation}
    For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
    \(c_{ua} := \cos(q_u + q_a)\). Likewise, \(s_u := \sin(q_u)\), 
    \(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\end{notation}

@@

A sufficient condition for regularity is when \(f\) is a function solely of \(p_u\),
because then \(\partial_{q_u}f = 0\) and (\ref{eqn:regularity-matrix-acrobot})
is strictly positive for all values of \(q_a\). 
This will be useful later, as we will not need to check regularity if we design
a function of the unactuated momentum.

@@

Let us now compute the constrained dynamics under
(\ref{eqn:acrobot-constraint}).
Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\), while
\[
    dh_{p_u} = \frac{-\bar{q}_a I}{1 + I^2 p_u^2}
    .
\]
Inserting these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
constraint manifold:
\[
    p_a(q_u,p_u) = \frac{
        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
    .
\]
The dynamics for \(p_u\) do not contain \(p_a\), so they remain unchanged.
The constrained dynamics for \(q_u\) are given by 
\begin{equation*}
    \dot{q}_u = e_1\tpose \Minv(q) \begin{bmatrix}
                    p_u \\ p_a(q_u,p_u)
                \end{bmatrix} %\\
    ,
\end{equation*}
%        &= \frac{1}{ml^2(2-c_a^2)} 
%            \left( p_u - 
%                (1+c_a)\frac{(1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
%                }{ml^2(3+2c_a)(1+I^2 p_u^2)}
%            \right) 
%        ,
%\end{align*}
which can be simplified to 
\begin{equation*}
    \dot{q_u} = \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }{ml^2(1+I^2 p_u^2)(3+2c_a)}
    .
\end{equation*}
Hence, the constrained dynamics for the acrobot under
(\ref{eqn:acrobot-constraint}) are
\begin{align}\label{eqn:acrobot-constrained-dynamics}
    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
            {ml^2(1+I^2 p_u^2)(3+2c_a)}
        , \\
    \dot{p}_u &= - m g l (2s_u + s_{ua})
    . \nonumber
\end{align}

@@

% TODO: This needs modification. The energy injection is provable for any system
% only during oscillations. For rotations, we need to do a numerical check of an
% integral that depends on m,g,l. If this integral is positive, then the system
% is gaining energy up to a point - it sems the integral decays to 0 so that you
% can reach a limiting speed, which means you don't get full energy injection
% everywhere on Gamma.
\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot (\ref{eqn:acrobot-hamiltonian}) constrained by
    (\ref{eqn:acrobot-constraint}) with \(\bar{q}_a \in ]0,1]\).
    There exists a small enough\footnotemark \(I^\star \in \R_{>0}\)
    such that the VNHC injects energy whenever \( I \in ]0,I^\star[\). 
    If \(I \in ]-I^\star,0[\), the VNHC dissipates energy.
\end{thm}
\footnotetext{In fact, we can be more precise than this. 
    Letting \(\epsilon > 0\), we need 
    \(I \leq \min\left\{I_1(\epsilon),I_2\right\}\) where
    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
    \right)^{-1/2}\)
    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
    See Apendix \ref{apx:I-bounds} for details.
}

@@

Letting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
One can use this to compute the solution \(x_1(t,\eta_0)\) of
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0
        .
    \end{cases}
\end{equation}
Finally, one can compute the first-order approximation
\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
term \(R(t,I)\) is \(O(I^2)\);
that is,
\[
    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
    .
\]
We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
    solution \(x(t,I,\eta_0)\) to (\ref{eqn:khalil-setup}) satisfies
    \[
        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_))\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

@@

\subsection*{Perturbation Analysis for Oscillations}
When \(I = 0\), the acrobot becomes a nominal pendulum that oscillates whenever the
nominal energy (\ref{eqn:acrobot-nominal-E}) is less than \(E(\pi,0)\).
Hence, the domain of oscillations on \(\Gamma\) is given by
\[
    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
We wish to find a transformation \(T(q_u,p_u)\) on \(D\) into
\((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
and \(\mu > 0\) is a pseudo-radius.
Since the energy of oscillation can be determined by the orbit's intersection
with the \(q_u\)-axis and \(q_u \in ]-\pi,\pi[\) on \(D\), 
we set \(\mu \in ]0,\pi[\) as in Figure \ref{fig:mu-intersection}.
The transformation we want is therefore a diffeomorphism of the form
\begin{align*}
    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi], \\
    (q_u, p_u) &\mapsto (\alpha,\mu)
    .
\end{align*}

@@

The energy level set corresponding to the intersection \((q_u,p_u) = (\mu,0)\)
is 
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = 3mgl(1- \cos(\mu))\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:oscillation-pu2}
    p_u^2 = 30m^2gl^3(\cos(q_u) - \cos(\mu))
    .
\end{equation}
On this level set, \(q_u\) ranges between \([-\mu,\mu]\) and can be uniquely
parameterized by \(q_u = \mu \cos(\alpha)\), where \(\alpha\) is our
pseudo-angle.
Substituting this into (\ref{eqn:oscillation-pu2}), we get
\[
    p_u^2 = 30m^2gl^3(\cos(\mu \cos(\alpha)) - \cos(\mu))
    .
\]
We want to find \(p_u\) as a function of \((\alpha,\mu)\); noting that we can
determine the sign of \(p_u\) from the sign of \(\sin(\alpha)\), we get the
(clockwise) parameterization
\begin{equation}\label{eqn:oscillation-pu}
    p_u = -\sign{\sin(\alpha)} \sqrt{30m^2gl^3 \left(\cos(\mu c_\alpha) - c_\mu\right)}
    ,
\end{equation}
which is, in fact, smooth.

We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\).
We need the inverse of this map to get our diffeomorphism \(T(q_u,p_u)\).
Notice from (\ref{eqn:oscillation-pu2}) that
\[
    \cos(\mu) = -\frac{p_u^2}{30m^2gl^3} + \cos(q_u) =: C_\mu(q_u,p_u)
    .
\]
Since \(\mu \in ]0,\pi]\), we can uniquely express \(\mu\) by
\begin{equation}\label{eqn:mu-qupu}
    \mu = \arccos\left(C_\mu(q_u,p_u)\right)
    .
\end{equation}
Next we need to find \(\alpha\). 
Recall that 
\[
    \cos(\alpha) = \frac{q_u}{\mu}
    ,
\] 
which means 
\[
    \sin(\alpha) = \pm \sqrt{1 - \frac{q_u^2}{\mu^2}}
    .
\]
Notice from (\ref{eqn:oscillation-pu}) that 
\(\sign{\sin(\alpha)} = -\sign{p_u}\).
We use these facts to deduce that
\begin{equation}\label{eqn:alpha-qupu}
    \alpha = \left.
        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
    ,
\end{equation}
and thus have our transformation \(T(q_u,p_u)\) into
\((\alpha,\mu)\)-coordinates on \(D\).

The acrobot's constrained dynamics in
\((\alpha,\mu)\)-coordinates can be computed by evaluating
\begin{align*}
    \dot{\mu} &= \left.\pdiff{\mu(q_u,p_u)}{q}_u\dot{q_u} +
         \pdiff{\mu(q_u,p_u)}{p_u}\dot{p}_u
         \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
    , \\
    \dot{\alpha} &= \left.\pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
        \pdiff{\alpha(q_u,p_u)}{p_u}\dot{p}_u
        \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
    .
\end{align*}
with \((\dot{q}_u,\dot{p}_u)\) the given by 
 (\ref{eqn:acrobot-constrained-dynamics}).

These dynamics are much too large to write out, so we simply denote them by
\begin{align}\label{eqn:acrobot-mu-dot}
    \dot{\mu} &= f_\mu(\alpha,\mu,I)
    ,\\
    \label{eqn:acrobot-alpha-dot}
    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
    .
\end{align}
The nominal system \((\alpha_n,\mu_n)\) comes from setting \(I = 0\).
MATLAB's symbolic toolbox evaluates the nominal dynamics as
\begin{align}\label{eqn:acrobot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-alpha-dot-nom}
    \dot{\alpha}_n &= \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu_n\cos(\alpha_n)) - \cos(\mu_n)}
            {\mu_n^2 \sin(\alpha_n)^2}}
    .
\end{align}
One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\) -- this
is expected, since the nominal system is a single pendulum.
By continuity of (\ref{eqn:acrobot-mu-dot}) with respect to \(I\), 
\(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).

@@

In the spirit of perturbation analysis, we expand our solution of
(\ref{eqn:muhat-dot}) into
\begin{equation}\label{eqn:acrobot-muhat-approx}
    \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
    + R(\alpha,I)
    ,
\end{equation}
where \(R(\alpha,I)\) is \(O(I^2)\).
From (\ref{eqn:khalil-perturbation-nominal}) we know that 
\(\hat{\mu}_0\) is the solution to the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
    ,
\end{cases}
\]
Since \(g(\alpha,\mu,0) = 0\) by 
(\ref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}), we find that
\(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).

>>>>

Since we are now working with a simple acrobot, 
we have \(l_{c_u} = l_{c_a} = l_u = l_a = l\) and
\(m_u = m_a = m\). 
On top of this, the moments of inertia \(J_u\) and \(J_a\) of the rods vanish.
Reducing
~\eqref{eqn:general-acrobot-inertia}-\ref{eqn:general-acrobot-potential}
yields the simplified inertia matrix \(D_s\) and potential function \(P_s\),
where
\begin{align}
    D_s(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    P_s(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    .
\end{align}
\begin{notation}
    For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
    \(c_{ua} := \cos(q_u + q_a)\). Likewise, \(s_u := \sin(q_u)\), 
    \(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\end{notation}

@@

A sufficient condition for regularity is when \(f\) is a function solely of \(p_u\),
because then \(\partial_{q_u}f = 0\) and ~\eqref{eqn:regularity-matrix-acrobot}
is strictly positive for all values of \(q_a\). 
This will be useful later, as we will not need to check regularity if we design
a function of the unactuated momentum.

@@

Let us now compute the constrained dynamics under
~\eqref{eqn:acrobot-constraint}.
Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\), while
\[
    dh_{p_u} = \frac{-\bar{q}_a I}{1 + I^2 p_u^2}
    .
\]
Inserting these to ~\eqref{eqn:g-qupu}, we get the solution for \(p_a\) on the
constraint manifold:
\[
    p_a(q_u,p_u) = \frac{
        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
    .
\]
The dynamics for \(p_u\) do not contain \(p_a\), so they remain unchanged.
The constrained dynamics for \(q_u\) are given by 
\begin{equation*}
    \dot{q}_u = e_1\tpose \Minv(q) \begin{bmatrix}
                    p_u \\ p_a(q_u,p_u)
                \end{bmatrix} %\\
    ,
\end{equation*}
%        &= \frac{1}{ml^2(2-c_a^2)} 
%            \left( p_u - 
%                (1+c_a)\frac{(1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
%                }{ml^2(3+2c_a)(1+I^2 p_u^2)}
%            \right) 
%        ,
%\end{align*}
which can be simplified to 
\begin{equation*}
    \dot{q_u} = \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }{ml^2(1+I^2 p_u^2)(3+2c_a)}
    .
\end{equation*}
Hence, the constrained dynamics for the acrobot under
~\eqref{eqn:acrobot-constraint} are
\begin{align}\label{eqn:acrobot-constrained-dynamics}
    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
            {ml^2(1+I^2 p_u^2)(3+2c_a)}
        , \\
    \dot{p}_u &= - m g l (2s_u + s_{ua})
    . \nonumber
\end{align}

@@

% TODO: This needs modification. The energy injection is provable for any system
% only during oscillations. For rotations, we need to do a numerical check of an
% integral that depends on m,g,l. If this integral is positive, then the system
% is gaining energy up to a point - it sems the integral decays to 0 so that you
% can reach a limiting speed, which means you don't get full energy injection
% everywhere on Gamma.
\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
    ~\eqref{eqn:acrobot-constraint} with \(\bar{q}_a \in ]0,1]\).
    There exists a small enough\footnotemark \(I^\star \in \R_{>0}\)
    such that the VNHC injects energy whenever \( I \in ]0,I^\star[\). 
    If \(I \in ]-I^\star,0[\), the VNHC dissipates energy.
\end{thm}
\footnotetext{In fact, we can be more precise than this. 
    Letting \(\epsilon > 0\), we need 
    \(I \leq \min\left\{I_1(\epsilon),I_2\right\}\) where
    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
    \right)^{-1/2}\)
    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
    See Apendix \ref{apx:I-bounds} for details.
}

@@

Letting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
One can use this to compute the solution \(x_1(t,\eta_0)\) of
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0
        .
    \end{cases}
\end{equation}
Finally, one can compute the first-order approximation
\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
term \(R(t,I)\) is \(O(I^2)\);
that is,
\[
    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
    .
\]
We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem ~\eqref{eqn:khalil-perturbation-nominal} has a
    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
    solution \(x(t,I,\eta_0)\) to ~\eqref{eqn:khalil-setup} satisfies
    \[
        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_))\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

@@

\subsection*{Perturbation Analysis for Oscillations}
When \(I = 0\), the acrobot becomes a nominal pendulum that oscillates whenever the
nominal energy ~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\).
Hence, the domain of oscillations on \(\Gamma\) is given by
\[
    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
We wish to find a transformation \(T(q_u,p_u)\) on \(D\) into
\((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
and \(\mu > 0\) is a pseudo-radius.
Since the energy of oscillation can be determined by the orbit's intersection
with the \(q_u\)-axis and \(q_u \in ]-\pi,\pi[\) on \(D\), 
we set \(\mu \in ]0,\pi[\) as in Figure \ref{fig:mu-intersection}.
The transformation we want is therefore a diffeomorphism of the form
\begin{align*}
    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi], \\
    (q_u, p_u) &\mapsto (\alpha,\mu)
    .
\end{align*}

@@

The energy level set corresponding to the intersection \((q_u,p_u) = (\mu,0)\)
is 
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = 3mgl(1- \cos(\mu))\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:oscillation-pu2}
    p_u^2 = 30m^2gl^3(\cos(q_u) - \cos(\mu))
    .
\end{equation}
On this level set, \(q_u\) ranges between \([-\mu,\mu]\) and can be uniquely
parameterized by \(q_u = \mu \cos(\alpha)\), where \(\alpha\) is our
pseudo-angle.
Substituting this into ~\eqref{eqn:oscillation-pu2}, we get
\[
    p_u^2 = 30m^2gl^3(\cos(\mu \cos(\alpha)) - \cos(\mu))
    .
\]
We want to find \(p_u\) as a function of \((\alpha,\mu)\); noting that we can
determine the sign of \(p_u\) from the sign of \(\sin(\alpha)\), we get the
(clockwise) parameterization
\begin{equation}\label{eqn:oscillation-pu}
    p_u = -\sign{\sin(\alpha)} \sqrt{30m^2gl^3 \left(\cos(\mu c_\alpha) - c_\mu\right)}
    ,
\end{equation}
which is, in fact, smooth.

We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\).
We need the inverse of this map to get our diffeomorphism \(T(q_u,p_u)\).
Notice from ~\eqref{eqn:oscillation-pu2} that
\[
    \cos(\mu) = -\frac{p_u^2}{30m^2gl^3} + \cos(q_u) =: C_\mu(q_u,p_u)
    .
\]
Since \(\mu \in ]0,\pi]\), we can uniquely express \(\mu\) by
\begin{equation}\label{eqn:mu-qupu}
    \mu = \arccos\left(C_\mu(q_u,p_u)\right)
    .
\end{equation}
Next we need to find \(\alpha\). 
Recall that 
\[
    \cos(\alpha) = \frac{q_u}{\mu}
    ,
\] 
which means 
\[
    \sin(\alpha) = \pm \sqrt{1 - \frac{q_u^2}{\mu^2}}
    .
\]
Notice from ~\eqref{eqn:oscillation-pu} that 
\(\sign{\sin(\alpha)} = -\sign{p_u}\).
We use these facts to deduce that
\begin{equation}\label{eqn:alpha-qupu}
    \alpha = \left.
        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
    ,
\end{equation}
and thus have our transformation \(T(q_u,p_u)\) into
\((\alpha,\mu)\)-coordinates on \(D\).

The acrobot's constrained dynamics in
\((\alpha,\mu)\)-coordinates can be computed by evaluating
\begin{align*}
    \dot{\mu} &= \left.\pdiff{\mu(q_u,p_u)}{q}_u\dot{q_u} +
         \pdiff{\mu(q_u,p_u)}{p_u}\dot{p}_u
         \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
    , \\
    \dot{\alpha} &= \left.\pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
        \pdiff{\alpha(q_u,p_u)}{p_u}\dot{p}_u
        \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
    .
\end{align*}
with \((\dot{q}_u,\dot{p}_u)\) the given by 
 ~\eqref{eqn:acrobot-constrained-dynamics}.

These dynamics are much too large to write out, so we simply denote them by
\begin{align}\label{eqn:acrobot-mu-dot}
    \dot{\mu} &= f_\mu(\alpha,\mu,I)
    ,\\
    \label{eqn:acrobot-alpha-dot}
    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
    .
\end{align}
The nominal system \((\alpha_n,\mu_n)\) comes from setting \(I = 0\).
MATLAB's symbolic toolbox evaluates the nominal dynamics as
\begin{align}\label{eqn:acrobot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-alpha-dot-nom}
    \dot{\alpha}_n &= \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu_n\cos(\alpha_n)) - \cos(\mu_n)}
            {\mu_n^2 \sin(\alpha_n)^2}}
    .
\end{align}
One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\) -- this
is expected, since the nominal system is a single pendulum.
By continuity of ~\eqref{eqn:acrobot-mu-dot} with respect to \(I\), 
\(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).

@@

In the spirit of perturbation analysis, we expand our solution of
~\eqref{eqn:muhat-dot} into
\begin{equation}\label{eqn:acrobot-muhat-approx}
    \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
    + R(\alpha,I)
    ,
\end{equation}
where \(R(\alpha,I)\) is \(O(I^2)\).
From ~\eqref{eqn:khalil-perturbation-nominal} we know that 
\(\hat{\mu}_0\) is the solution to the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
    ,
\end{cases}
\]
Since \(g(\alpha,\mu,0) = 0\) by 
~\eqref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}, we find that
\(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).

Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to ~\eqref{eqn:acrobot-mu1-dot} is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]
Notice that \(\hat{b}(\alpha,\mu_0)\) is both even and 
\(\pi\)-periodic in \(\alpha\), so that the following properties hold:
\begin{equation}\label{eqn:bhat-integral-properties}
    \begin{split}
    \int \limits_0 ^ {N \pi} \hat{b}(\sigma,\mu_0) d\sigma
    &= N \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma \, 
    \left(\forall N \in \mathbb{Z}_{> 0}\right)
    , \\
    \int \limits_{-\pi}^0 \hat{b}(\sigma,\mu_0) d\sigma
    &= \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    .
    \end{split}
\end{equation}
Suppose we start the system at \((\alpha,\mu) = (0,\mu_0)\). 
Then the new pseudo-radius after a half orbit where \(\alpha\) goes from \(0\)
to \(\pi\) (that is, when the orbit hits the negative \(q_u\)-axis)
is \(\rho = \hat{\mu}(\pi,\mu_0)\).
Hence, the change in pseudo-radius is \(\rho - \mu_0 = I\hat{\mu}_1(\pi,\mu_0)\).
Now look at the full orbit where \(\alpha\) goes from \(0\) to \(2\pi\). 
This lands on the positive \(q_u\)-axis with pseudo-radius
\(\tilde{\rho} = \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
As we know from ~\eqref{eqn:bhat-integral-properties}, we have
\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\). 
This means the change in pseudo-radius after one entire oscillation is 
\(\tilde{\rho} - \mu_0 = 2I\hat{\mu}1(\pi,\mu_0) = 2 (\rho - \mu_0)\).
Hence, if we can show that \(\hat{\mu}_1(\pi,\cdot)\) is always positive, we
will guarantee that the pseudo-radius increases after every half oscillation.
This in turn guarantees \(q_u\) is increasing after every half oscillation, so
that eventually we will reach a rotation.

To show that the first-order approximation ~\eqref{eqn:acrobot-muhat-approx} is
gaining energy, we define the Poincar\'{e} section
\(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).
